---
title: "Data I/O Lab, Part 1"
output: html_document
---

RUN THIS BEFORE YOU START TO CLEAR ANY VARIABLES/OBJECTS FROM THE R ENVIRONMENT THAT COULD INTERFERE WITH YOUR WORK IN THIS ACTIVITY.
```{r}
rm(list=ls())
```


This lab uses different versions of the `iris` dataset, which you can find in the data/iris subfolder. (This dataset is also included as an example to play with in standard R installations. You can learn more about it by running `?iris` in the console.)

0. Load necessary libraries

```{r} 
library(tidyverse)
library(readxl)
library(here)
```

1.  Explore the different versions of the iris data files to see what type of data delimiter each one employs. In the file explorer panel in RStudio, navigate to the data/iris directory and then click on each file one at a time. For most of the files, clicking on it will bring up a dialog box with two options: "View File" and "Import Dataset...". Click on "View File". 

2.  Create path variable for the iris data subdirectory using `here()`

```{r}
iris_data_dir <- here("data/iris//") 
# For some reason, you only need to have a double slash at the end of the directory
# string, not in the middle. Not sure why at this point in my R life...
```

3. Read in the `iris_q1.csv` dataset into the `iris_q1` R object. What delimiter separates columns?

```{r}
iris_q1 <- read_csv(paste0(iris_data_dir, "iris_q1.csv"))
head(iris_q1)
```

4. Read in the `iris_q2.tsv` dataset into the `iris_q2` R object. What delimiter separates columns?

```{r}
iris_q2 <- read_delim(paste0(iris_data_dir, "iris_q2.tsv"), delim = "\t")
head(iris_q2)
```

5. Read in the `iris_q3.txt` dataset into the `iris_q3` R object. What delimiter separates columns?

```{r}
iris_q3 <- read_delim(paste0(iris_data_dir, "iris_q3.txt"), delim = " ")
head(iris_q3)
```

6. Read in the `iris_q4.csv` dataset into the `iris_q4` R object. What delimiter separates columns? [hint: file extension is ambiguous]

```{r}
iris_q4 <- read_delim(paste0(iris_data_dir, "iris_q4.csv"), delim = ":")
head(iris_q4)
```

7. Read in the `iris_q5.csv` dataset into the `iris_q5` R object. Some entries are supposed to be missing, e.g. have the value `NA` in the data.  

```{r}
iris_q5 <- read_csv(paste0(iris_data_dir, "iris_q5.csv"), na = "-999")
head(iris_q5)
```

8. Read in the `iris_q6.xlsx` dataset into the `iris_q6` R object. How many rows are in the dataset?

```{r}
iris_q6 <- read_excel(paste0(iris_data_dir, "iris_q6.xlsx"), sheet= 2)
dim(iris_q6)
```
