---
title: "Data Cleaning lab key"
output: html_document
---

RUN THIS BEFORE YOU START TO CLEAR ANY VARIABLES/OBJECTS FROM THE R ENVIRONMENT THAT COULD INTERFERE WITH YOUR WORK IN THIS ACTIVITY.
```{r}
rm(list=ls())
```


```{r}
library(tidyverse)
library(here)
```

1. Read in the UFO dataset (used in the Data IO lectures) as an R object called `ufo`.
You can open it from the data/ufo subdirectory, or you can read directly from the web here: https://raw.githubusercontent.com/SISBID/Module1/gh-pages/data/ufo/ufo_data_complete.csv . 
You can ignore the "problems" with some rows. 

```{r}
ufo_data_dir <- here("data/ufo//")

ufo <- read_csv(paste0(ufo_data_dir, "ufo_data_complete.csv"))

head(ufo)
```

2. Clean up the column/variable names of the `ufo` dataset to remove spaces and non-alphanumeric characters. Load the `janitor` library (install it first if neded!) then run `clean_names()` on the dataframe. 

```{r}
library(janitor)
ufo <- clean_names(ufo)
head(ufo)
```

3. How many UFO sightings were reported on a time scale of minutes, specifically using the `duration (hours/min)` originally-named column? (hint: explore the various entries in this column and then use a `stringr` function to match a flexible pattern. you can ignore observations like "1/2 hour" and similar ones that don't contain some version of the word "minutes")

```{r}
str_subset(ufo$duration_hours_min, "min") %>% length()
```

4. How accurate is the (formerly-named) `duration (seconds)` column? I.e. how many of the above minutes-scale observations have durations less than 60 seconds or greater than 3600 seconds? 

```{r}
sub <- ufo %>% filter(str_detect(duration_hours_min, "min"))
mean(sub$duration_seconds < 60 | sub$duration_seconds > 3600)
```

5. How many ufo sighting cities end in (a) "field", (b) "ton" and (c) "port"?


```{r}
str_subset(ufo$city, "field$") %>% length() # A
str_subset(ufo$city, "ton$") %>% length() # B
str_subset(ufo$city, "port$") %>% length() # B
```

6. Drop countries that contain ufo sightings/rows where the listed states are possibly indicative of errors. How many countries are realistically represented in these data? (hint: look at cross-tabulations of state and country)

```{r}
table(ufo$state, ufo$country)
ufo <- ufo %>% filter(country %in% c("ca", "us"))
```
