---
title: "Data Summarization Lab"
output: html_document
editor_options: 
  chunk_output_type: console
---

RUN THIS BEFORE YOU START TO CLEAR ANY VARIABLES/OBJECTS FROM THE R ENVIRONMENT THAT COULD INTERFERE WITH YOUR WORK IN THIS ACTIVITY.
```{r}
rm(list=ls())
```


# Data used

Circulator Lanes Dataset: the data is from https://data.baltimorecity.gov/Transportation/Charm-City-Circulator-Ridership/wwvu-583r

You can download CSV from the site above to your current working directory, or you can load it directly from the data/ folder in this repo. Run the code chunk below to load the data set into an R object called `circ`.

```{r}
library(tidyverse) # to use `read_csv()`
library(here) # to use relative file paths

data_dir <- here("data//") 
# to create a string variable that contains the directory path to 
# the data/ subdirectory on your computer. R will interpret the double
# backslash after "data" as a single backslash when it creates the string

circ <- read_csv(paste0(data_dir, "Charm_City_Circulator_Ridership.csv"))
```

What is contained in the object `data_dir`? Run the code below to find out:

```{r}
data_dir
```



1. How days are in the data set?  You can assume each observation/row is a different day
```{r q1}

```

2. What is the total (sum) number boardings on the green bus?

```{r q2}

```

3.  How many days are missing daily ridership?  Use `is.na` and either table, count?
   
```{r q3}

```


4. Which day of the week has the most average ridership (using `daily`)?  How many days goes are observed for those days? Group the data.

```{r q4}

```



5. (a) Numerically [hint: `quantile()`] and 
		(b) graphically [hint: `qplot(, data = circ)`]
		 describe the distribution of `orangeBoardings`
```{r q6}

```

6. Then describe as above, after stratifying by day of the week.
```{r q7}

```

7.  Describe the following output and then flip the `x` and `y` variables:
```{r}
qplot(x = orangeBoardings, y = day, data = circ, geom = "boxplot")
```


