---
title: "Data Reshaping Lab Key"
author: "Data Wrangling in R"
output: html_document
---

RUN THIS BEFORE YOU START TO CLEAR ANY VARIABLES/OBJECTS FROM THE R ENVIRONMENT THAT COULD INTERFERE WITH YOUR WORK IN THIS ACTIVITY.
```{r}
rm(list=ls())
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(here)
```

1.  This lab uses the the Charm City Circulator Dataset, the same data set from Labs 2 and 4. Read the data set into an R object called `wide`.

```{r}

```

2.  Reshape `wide` using `pivot_long` Call this data `long` Assign "route_info" to `names_to`, "n_people" to `values_to`. Make sure to pivot all columns **but** `daily`, `day`, `date`, using things like `-day`. Table `long$day`

```{r}

```

Run the code chunk below to compare the structure of `wide` and `long`.

```{r}
wide
long
```

3.  Use the `separate()` function to separate the `date` column into ("month", "day", "year") using the "/" character. Make sure you don't remove the `date` column. Call this `long2`. Create a table of the day column. Where did the original `day` column go?

```{r}


```

Run the code chunk below to compare the structure of `long` and `long2`.

```{r}
long
long2
```

4.  Create a new variable in `long2` calling it `newdate` using the `unite` function. Unite the year, month, day, columns using a hyphen (`-`) and then `table` the new column `newdate`

```{r}

```

5.  Back to `long`. Add the "\_" character to the `route_info` values between the route name ("orange", "purple", "green", "banner") and the info ("Boardings", "Alightings", "Average") using `mutate` and `str_replace`.

```{r, error = TRUE}

```

6.  Separate the `route_info` column from `long` (not `long2`) into ("route", "type") using a sep of "\_"

```{r}

```

7.  Reshape `long` (not `long2`) using `pivot_wider`. Take the names from `route` and the values from `n_people`. Call this `wide_by_route`

```{r}

```
